---
title: "middlesex vagrants"
author: "Sharon Howard"
date: "3/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# packages
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(vcd)
library(vcdExtra)
```


```{r}
#Middlesex Vagrants v1.1 

# read in the data
mv <- read.csv("data/mv_20180328.csv", header=TRUE)

# add a column for group type (n=1 single, n>1 group)
mv <- mv %>%
  mutate(group_type =  ifelse(NumberofPeopleinGroup==1, "single", "group") )

# I want only lead/single vagrants who have known gender - filter out dependents and any [unknown] gender 

mv_lead_gender <- mv %>%
  filter(PersonType != "Dependent", GenderofLeadVagrant !="[unknown]")

# aggregate for gender
mv_gen <- mv_lead_gender %>% 
         select(GenderofLeadVagrant) %>%
         dplyr::group_by(GenderofLeadVagrant) %>%
         dplyr::summarise(n_g = n())

# aggregrate by gender and group type
mv_gen_grp <- mv_lead_gender %>% 
         select(GenderofLeadVagrant, group_type) %>%
         dplyr::group_by(GenderofLeadVagrant, group_type) %>%
         dplyr::summarise(n_g = n())

# aggregate by gender and year of session
mv_gen_year <- mv_lead_gender %>%
         select(SessionStartYear,GenderofLeadVagrant) %>%
         dplyr::group_by(SessionStartYear,GenderofLeadVagrant) %>%
         dplyr::summarise(n_year = n())

# aggregate by gender and settlement country
mv_gen_ctry <-  mv_lead_gender %>%
         dplyr::filter(grepl("England|Scotlan|Ireland|Wales", SettlementCountry)) %>%
         select(GenderofLeadVagrant,SettlementCountry) %>%
         dplyr::group_by(GenderofLeadVagrant,SettlementCountry) %>%
         dplyr::summarise(n_g = n())

# aggregate gender, settlement country, type
mv_gen_ctry_grp <- mv %>%
         filter(PersonType != "Dependent", GenderofLeadVagrant !="[unknown]", grepl("England|Scotlan|Ireland|Wales", SettlementCountry)) %>%
         select(GenderofLeadVagrant, SettlementCountry, group_type) %>%
         dplyr::group_by(group_type,SettlementCountry,GenderofLeadVagrant) %>%
         dplyr::summarise(n_gs = n())

# same thing with a few tweaks for the mosaic plot
mv_lead_mos <-  mv_lead_gender %>% 
  dplyr::filter(PersonType != "Dependent", GenderofLeadVagrant !="[unknown]", grepl("England|Scotlan|Ireland|Wales", SettlementCountry) ) %>%  
  droplevels %>%  # this *properly* gets rid of the filtered stuff for the mosaic plot!
  select(GenderofLeadVagrant, SettlementCountry, group_type)  %>%
  rename(gender=GenderofLeadVagrant, country=SettlementCountry, type=group_type)
```



Compare overall numbers male/female group leader and single vagrants

```{r}
ggplot(mv_gen, aes(y=n_g, x=GenderofLeadVagrant)) +
  geom_bar(stat = "identity", aes(fill=GenderofLeadVagrant)) +
  labs(x="gender", y="number of vagrants", fill="", title="Gender of Middlesex Vagrants 1777-1786")

#ggsave(  "mv_gender.png",  width = 6,  height = 4,  dpi = 100)
```


Grouped

```{r}
ggplot(mv_gen_year, aes(x=SessionStartYear, y=n_year, fill=GenderofLeadVagrant, group=GenderofLeadVagrant)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_x_continuous(breaks = c(1777, 1778, 1779, 1780, 1781, 1782, 1783, 1784,1785,1786)) +
  labs(fill="gender", x="year", y="number of vagrants", title="Gender of Middlesex Vagrants by year, 1777-1786")

#ggsave( "mv_gender_year.png",  width = 6,  height = 4,  dpi = 100)
```

I think it's possibly less immediately obvious in the stacked chart that the m:f ratio flips after 1782 as numbers of male vagrants start to go up massively? Prefer above anyway.

```{r}
ggplot(mv_gen_year, aes(x=SessionStartYear, y=n_year, fill=GenderofLeadVagrant)) +
  geom_bar(stat="identity") +
  scale_x_continuous(breaks = c(1777, 1778, 1779, 1780, 1781, 1782, 1783, 1784,1785,1786)) +
  labs(fill="gender", x="year", y="number of vagrants")
```



The majority of vagrants are lone individuals not in groups, but women are more likely to be in groups

```{r}

ggplot(mv_gen_grp , aes(x=GenderofLeadVagrant, y=n_g, fill=group_type)) +
  geom_bar(stat="identity", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
 labs(x="gender", y="% of vagrants", fill="vagrant type", title="Type of Middlesex Vagrants, by gender")


#ggsave( "mv_type_gender.png", width = 6, height = 4, dpi = 100)
 
```
For single vagrants, the gender split is not far off 50/50, but slightly higher proportion are male. But again can see that a considerably higher proportion of women than men are in groups.

```{r}

ggplot(mv_gen_grp, aes(x=group_type, y=n_g, fill=GenderofLeadVagrant)) +
  geom_bar(stat="identity", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(fill="gender", y="% of vagrants", x="vagrant type", title="Gender of Middlesex Vagrants, by type")
 

#ggsave("mv_gender_type.png", width = 6, height = 4, dpi = 100)
```


```{r}
ggplot(mv_gen_ctry, aes(x=GenderofLeadVagrant, y=n_g, fill=SettlementCountry)
) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(fill="settlement\ncountry", x="gender", y="% of vagrants", title="Gender and settlement country of Middlesex vagrants")


#ggsave( "mv_gender_country.png", width = 6, height = 4, dpi = 100)
  
```

```{r}
ggplot(mv_gen_ctry, aes(fill=GenderofLeadVagrant, y=n_g, x=SettlementCountry)
) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(x="settlement country", fill="gender", y="% of vagrants")
  
```


```{r}
ggplot(mv_gen_ctry_grp, aes(x=group_type, y=n_gs, fill=SettlementCountry)
) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~GenderofLeadVagrant) +
  scale_y_continuous(labels = percent_format())  +
  labs(fill="settlement\ncountry", x="type", y="% of vagrants", title="Compare gender and country of vagrants by type")
  

#ggsave("mv_gender_country_type.png", width = 6, height = 4, dpi = 100)
```

```{r}
ggplot(mv_gen_ctry_grp , aes(fill=group_type, y=n_gs, x=SettlementCountry)
) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~GenderofLeadVagrant) +
  scale_y_continuous(labels = percent_format()) +
  labs(x="settlement country", fill="group\ntype", y="% of vagrants")
  
```




```{r}
#png(filename = "mv_mosaic2.png", width = 1000, height = 600, res = 100)

mv_df <- data.frame(mv_lead_mos)

mv_mosaic <- mosaic(~ gender + country + type , mv_df, main="Mosaic plot of vagrants' gender/country/type", color = TRUE, shade=TRUE, labeling=labeling_border( rot_labels = c(40,90,0,90), gp_labels = gpar(fontsize = 11) ) )

#dev.off()

```





```{r}
kable(mv_gen_ctry_grp)
```

