---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---



```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```


```{r}

wbspse <- read.csv("data/wbspse_alldates_20180306.csv")


# convert wbig to long format, 1960-2009 only
wbspse_long_all_ind <- wbspse %>%
    select(IndicatorCode, IndicatorName, CountryName, CountryCode, X1960, X1961, X1962, X1963, X1964, X1965, X1966, X1967, X1968, X1969, X1970, X1971, X1972, X1973, X1974, X1975, X1976, X1977, X1978, X1979, X1980, X1981, X1982, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008, X2009) %>%
    gather(year1, val, X1960, X1961, X1962, X1963, X1964, X1965, X1966, X1967, X1968, X1969, X1970, X1971, X1972, X1973, X1974, X1975, X1976, X1977, X1978, X1979, X1980, X1981, X1982, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008, X2009) %>%
    mutate(year = as.integer(substr(year1, 2,5)), 
           decade = paste0(substr(year1, 2,4), '0s') )


wbsp1960 <- read.csv("data/wbsp_1960_20180306.csv",
                   header=TRUE)

# convert wbsp1960 to long format, 1960-2009 only
wbsp1960_long_all_ind <- wbsp1960 %>%
    select(IndicatorCode, IndicatorName, CountryName, CountryCode, X1960, X1961, X1962, X1963, X1964, X1965, X1966, X1967, X1968, X1969, X1970, X1971, X1972, X1973, X1974, X1975, X1976, X1977, X1978, X1979, X1980, X1981, X1982, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008, X2009) %>%
    gather(year1, val, X1960, X1961, X1962, X1963, X1964, X1965, X1966, X1967, X1968, X1969, X1970, X1971, X1972, X1973, X1974, X1975, X1976, X1977, X1978, X1979, X1980, X1981, X1982, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008, X2009) %>%
    mutate(year = as.integer(substr(year1, 2,5)), 
           decade = paste0(substr(year1, 2,4), '0s') )



```

```{r}

palette <- c("#D55E00","#0072B2", "#56B4E9", "#999999", "#009E73", "#F0E442", "black")

filter_fertility <- "SP.DYN.TFRT.IN"
filter_fertility_teen <- "SP.ADO.TFRT"

filter_life <- "SP.DYN.LE00.(FE|MA).IN"

filter_primary <- "SE.PRM.ENRL.FE.ZS"
filter_secondary <- "SE.SEC.ENRL.FE.ZS"
filter_primary_teachers <- "SE.PRM.TCHR.FE.ZS"

filter_secondary_teachers <- "SE.SEC.TCHR.FE.ZS"
filter_literacy <- "SE.ADT.LITR.(FE|MA).ZS"
filter_literacy_youth <- "SE.ADT.1524.LT.FE.ZS|SE.ADT.1524.LT.MA.ZS"

filter_income <- "HIC|LIC|LMC|MIC|UMC|WLD" 

filter_six <- "TUN|PAK|NGA|JAM|NOR|FJI"

filter_sixw <- "TUN|PAK|NGA|JAM|NOR|FJI|WLD"

```



```{r}
ggplot(wbsp1960_long_all_ind %>%
  filter(grepl(filter_income, CountryCode), grepl(filter_life, IndicatorCode) ),
  aes(year, val, colour=IndicatorCode)
  ) + 
  geom_line() +
  facet_wrap(~CountryName) +
  scale_colour_manual(values=palette, labels=c("SP.DYN.LE00.FE.IN"="female", "SP.DYN.LE00.MA.IN"="male"), name="Gender") +
  labs(title="Life expectancy by country income groups, 1960-2009")
```



```{r}

ggplot(wbsp1960_long_all_ind %>%
  filter(grepl(filter_six, CountryCode), grepl(filter_life, IndicatorCode) ),
  aes(year, val, colour=IndicatorCode)
  ) + 
  geom_line() +
  facet_wrap(~CountryName) +
  scale_colour_manual(values=palette, labels=c("SP.DYN.LE00.FE.IN"="female", "SP.DYN.LE00.MA.IN"="male"), name="Gender") +
  labs(title="Life expectancy in six countries, 1960-2009")
```


```{r}
ggplot(wbsp1960_long_all_ind %>%
      filter(grepl(filter_six, CountryCode), grepl(filter_life, IndicatorCode))
       , aes(year, val)) +
  geom_point(aes(colour = CountryName, shape=IndicatorCode) )  +
  labs(y="years", title="Life expectancy in six countries") +
  scale_colour_manual(values=palette, name="Countries" ) +
  scale_shape_discrete(labels=c("SP.DYN.LE00.FE.IN"="female", "SP.DYN.LE00.MA.IN"="male"), name="Gender",solid = FALSE) 
```


Fertility rate, total (births per woman)

```{r}
ggplot(wbsp1960_long_all_ind %>%
      filter(grepl(filter_income, CountryCode), grepl(filter_fertility, IndicatorCode))
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  scale_colour_manual(values=palette, name="Countries", breaks=c("High income", "Upper middle income", "Middle income", "Lower middle income", "Low & middle income", "Low income", "World")) +
  labs(y="births per woman", title="Fertility rates by country income groups")
```

Adolescent fertility rate (births per 1,000 women ages 15-19)

```{r}
ggplot(wbsp1960_long_all_ind %>%
      filter(grepl(filter_six, CountryCode), grepl(filter_fertility, IndicatorCode)), aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="births per woman", title="Fertility rates in six countries") +
  scale_colour_manual(values=palette, name="Countries" )
```


Primary education, pupils (% female)

```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_income, CountryCode), grepl(filter_primary, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of pupils", title="Primary education of girls by country income groups") +
  scale_colour_manual(values=palette, name="Countries", breaks=c("High income", "Upper middle income", "Middle income", "Lower middle income", "Low & middle income", "Low income", "World") )
```


```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_sixw, CountryCode), grepl(filter_primary, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of pupils", title="Primary education of girls in six countries") +
  scale_colour_manual(values=palette, name="Countries" )

```




Secondary education, pupils (% female)

```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_income, CountryCode), grepl(filter_secondary, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of pupils", title="Secondary education of girls by country income groups") +
  scale_colour_manual(values=palette, name="Countries", breaks=c("High income", "Upper middle income", "Middle income", "Lower middle income", "Low & middle income", "Low income", "World") )
```

```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_sixw, CountryCode), grepl(filter_secondary, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of pupils", title="Secondary education of girls in six countries") +
  scale_colour_manual(values=palette, name="Countries")
```



```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_income, CountryCode), grepl(filter_primary_teachers, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of teachers", title="Female primary school teachers") +
  scale_colour_manual(values=palette, name="Countries", breaks=c("High income", "Upper middle income", "Middle income", "Lower middle income", "Low & middle income", "Low income", "World") )
```

```{r}
ggplot(wbspse_long_all_ind %>%
      filter(grepl(filter_sixw, CountryCode), grepl(filter_primary_teachers, IndicatorCode), year>1969)
       , aes(year, val, colour = CountryName)) +
  geom_line() +
  labs(y="% of teachers", title="Female primary school teachers in six countries") +
  scale_colour_manual(values=palette, name="Countries" )
```





When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

